---
title: "Activity Evaluation"
author: "Zine Eddine"
date: '2022-07-04'
output: html_document
---

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project We will predict the manner in which they did the exercises.

## Loading The Data

```{r, cache=TRUE}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

## Removing identification columns

```{r, cache=TRUE}
training <- training[, 6:ncol(training)]
testing <- testing[, 6:ncol(testing)]
```

## Data Exploration

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(caret)
library(lubridate)
library(randomForest)
```

### removing columns that has too much missing values or empty values.

```{r, cache=TRUE , message=FALSE}
col.na <- apply(training, 2, function(x){sum(is.na(x))})
col.empty <- apply(training, 2, function(x){sum(x == "", na.rm = TRUE)})
col.notna <- names( col.na[col.na < 300])
col.notempty <- names( col.empty[col.empty < 300])
col.keep <- intersect(col.notna, col.notempty)


df_train <- training %>% select(col.keep)
df_test <- testing %>% select(setdiff(col.keep, c("classe")))
```

We need to convert the class column to a factor.

```{r}
df_train <- df_train %>%
            mutate(classe = as.factor(classe))

```


Because the data is too big I will use just 3000 rows.

```{r, cache=TRUE}
tdf_train <- df_train %>% 
             group_by(classe) %>%
             sample_n(600)

tdf_train <- tdf_train[sample(1:nrow(tdf_train)), ]
```



## Training The Models

We will use random forest because it has a very good accuracy.

### Random Forest


```{r, cache=TRUE}
set.seed(21432)

tr.control <- trainControl(method = "cv",
                           number = 10,
                           allowParallel = TRUE)

fit.rf <- train(classe ~ ., method = "rf", trControl = tr.control, data = tdf_train)
fit.rf
```

The model Accuracy is 0.96 which is ok for this kind of datasets.

```{r}
preds <- predict(fit.rf, newdata = df_test)
preds
```

The predictions in this RMD file got 100% on the quiz test.

